---
title: "coding_challenge_social_influence"
format: html
---

### building the network

loading the data

```{r}
url1 <- "https://github.com/JeffreyAlanSmith/Integrated_Network_Science/raw/master/data/class237_sem1_edgelist.txt"

sem1_edgelist <- read.table(file = url1, header = T)

url2 <- "https://github.com/JeffreyAlanSmith/Integrated_Network_Science/raw/master/data/class237_sem2_edgelist.txt"

sem2_edgelist <- read.table(file = url2, header = T)

url3 <- "https://github.com/JeffreyAlanSmith/Integrated_Network_Science/raw/master/data/class237_attributes.txt"

attributes <- read.table(file = url3, header = T)
```

loading packages

```{r}
library(igraph) 
library(reshape)
library(car)
```

filtering to only students who were there both semesters

```{r}
ids_sem1 <- attributes$ids[attributes$sem_id == 1]
ids_sem2 <- attributes$ids[attributes$sem_id == 2]

ids_keep <- ids_sem1[ids_sem1 %in% ids_sem2]
```

```{r}
attributes <- attributes[attributes$ids %in% ids_keep, ]

send_in_sem1 <- sem1_edgelist[,1] %in% ids_keep
rec_in_sem1 <- sem1_edgelist[,2] %in% ids_keep
sem1_edgelist <- sem1_edgelist[send_in_sem1 & rec_in_sem1, ]

send_in_sem2 <- sem2_edgelist[,1] %in% ids_keep
rec_in_sem2 <- sem2_edgelist[,2] %in% ids_keep
sem2_edgelist <- sem2_edgelist[send_in_sem2 & rec_in_sem2, ]
```

```{r}
sem1_attributes <- attributes[attributes$sem_id == 1, ]
sem2_attributes <- attributes[attributes$sem_id == 2, ]
```

creating the networks for each semester

```{r}
sem1_net <- graph_from_data_frame(d = sem1_edgelist, directed = T,
                                  vertices = sem1_attributes)

sem2_net <- graph_from_data_frame(d = sem2_edgelist, directed = T,
                                  vertices = sem2_attributes)
```

visualizing networks; colored by affect toward subject 

```{r}
cols_sem1 <- recode(sem1_attributes$like_subject, as.factor = F,
                    "1 = 'blue'; 2 = 'light blue'; 3 = 'pink'; 
                     4 ='red'; NA = 'black'")

cols_sem2 <- recode(sem2_attributes$like_subject, as.factor = F,
                    "1 = 'blue'; 2 = 'light blue'; 3 = 'pink'; 
                     4 = 'red'; NA = 'black'")
```

```{r}
layout <- layout_with_fr(sem1_net)

par(mfrow = c(1, 2))

plot(sem1_net, vertex.label = NA, layout = layout, edge.arrow.size = .3, 
     edge.arrow.width = 1, edge.color="light gray",
     vertex.frame.color = NA, vertex.color = cols_sem1) 

plot(sem2_net, vertex.label = NA, layout = layout, edge.arrow.size = .3,
     edge.arrow.width = 1, edge.color = "light gray",
     vertex.frame.color = NA, vertex.color = cols_sem2) 

# obvious changes in opinion
```

### SIENA model setup

```{r}
library(RSiena)
```

changing data to matrix form

```{r}
sem1_matrix <- as_adjacency_matrix(graph = sem1_net, sparse = F)
sem2_matrix <- as_adjacency_matrix(graph = sem2_net, sparse = F)

table(rownames(sem1_matrix) == rownames(sem2_matrix))
```

preliminary look at changes in ties between the semsters

```{r}
tab_change <- table(sem1_matrix, sem2_matrix)
tab_change

# 42 stable ties, 58 new, 47 dropped

42 / (58 + 47 + 42)

# similarity with Jaccard index; 28.6% of ties were present in both semesters
```

```{r}
table(sem1_attributes$ids == sem2_attributes$ids)
```

changes matrices to arrays

```{r}
net_size <- nrow(sem1_matrix)
net_size

sem12_array <- array(c(sem1_matrix, sem2_matrix), 
                     dim = c(net_size, net_size, 2))
```

creating tie status as the dependent variable

```{r}
networks <- sienaDependent(netarray = sem12_array, type = "oneMode") 
```

creating subject opinion as dependent variable

```{r}
like_subject_matrix <- as.matrix(cbind(sem1_attributes$like_subject,
                                       sem2_attributes$like_subject))

like_subject <- sienaDependent(netarray = like_subject_matrix, 
                               type = "behavior") 
```

creating covariate (opinion of teacher)

```{r}
like_teacher <- coCovar(sem1_attributes$like_teacher)
```

creating the SIENA object for the model

```{r}
net_behavior_covar <- sienaDataCreate(networks, like_subject, like_teacher)
net_behavior_covar

# can see basic stats like density and range of variables
```

### model specification

base model

```{r}
siena_effects <- getEffects(net_behavior_covar)
siena_effects
```

behavioral model

```{r}
siena_effects <- includeEffects(siena_effects, avSim, 
                                interaction1 = "networks",
                                name = "like_subject")  # peer influence/general opinion convergence


siena_effects <- includeEffects(siena_effects, indeg, 
                                interaction1 = "networks",
                                name = "like_subject") # effect of in-degree on subject opinion


siena_effects <- includeEffects(siena_effects, effFrom,
                                interaction1 = "like_teacher", 
                                name = "like_subject") # individual-level covariates (liking teacher)
```

network model

```{r}
siena_effects <- includeEffects(siena_effects, simX, 
                                 interaction1 = "like_subject", 
                                 name = "networks") # selection based on shared opinion


siena_effects <- includeEffects(siena_effects, egoX, 
                                 interaction1 = "like_subject", 
                                 name = "networks") # effect of subject opinion on out-degree


siena_effects <- includeEffects(siena_effects, altX, 
                                 interaction1 = "like_subject", 
                                 name = "networks") # effect of subject opinion on in-degree


siena_effects <- includeEffects(siena_effects, transTrip, name = "networks") # transitivity
```

### model estimation

```{r}
input_options <- sienaAlgorithmCreate(projname = "class_model", 
                                      MaxDegree = c(networks = 5), 
                                      seed = 30000) 
```

```{r}
mod1 <- siena07(alg = input_options, data = net_behavior_covar, 
                effects = siena_effects, useCluster = TRUE,
                nbrNodes = 2) 

mod1

# convergence under 0.25 which is good
# large standard errors; no significant results
```

trying a simpler model with less terms 

```{r}
siena_effects <- includeEffects(siena_effects, indeg, 
                                interaction1 = "networks",
                                name = "like_subject", include = FALSE) 

siena_effects <- includeEffects(siena_effects, effFrom,
                                interaction1 = "like_teacher", 
                                name = "like_subject", include = FALSE) 

# also increasing number of iterations
input_options <- sienaAlgorithmCreate(projname = "class_model", 
                                      MaxDegree = c(networks = 5), 
                                      n3 = 3000, seed = 30000)
```

```{r}
mod2 <- siena07(alg = input_options, data = net_behavior_covar, 
                effects = siena_effects, returnDeps = TRUE,
                useCluster = TRUE, nbrNodes = 2) 

mod2

# convergence still good
# still not significant results; likely due to small size of network and only 2 time periods
```

### checking model fit

in-degree

```{r}
gof1 <- sienaGOF(object = mod2, 
                 auxiliaryFunction = IndegreeDistribution, 
                 verbose = TRUE, join = TRUE, varName = "networks") 

plot(gof1)

# looks good
```

triad census

```{r}
gof2 <- sienaGOF(object = mod2, auxiliaryFunction = TriadCensus, 
                 verbose = TRUE, join = TRUE, varName = "networks")

plot(gof2, center = TRUE, scale = TRUE)   

# also looks good
```

behavior/opinion distribution

```{r}
gof_behaviour <- sienaGOF(object = mod2, 
                          auxiliaryFunction = BehaviorDistribution,
                          varName = "like_subject")

plot(gof_behaviour)

# also okay
```

### SIENA with 3 time periods

reading in the data

```{r}
url4 <- "https://github.com/JeffreyAlanSmith/Integrated_Network_Science/raw/master/data/Glasgow-friendship.RData"

load(url(description = url4))

url5 <- "https://github.com/JeffreyAlanSmith/Integrated_Network_Science/raw/master/data/Glasgow-substances.RData"

load(url(description = url5))

url6 <- "https://github.com/JeffreyAlanSmith/Integrated_Network_Science/raw/master/data/Glasgow-demographic.RData"

load(url(description = url6))

# friendship and alcohol use
```

removing value from ties

```{r}
friendship.1[friendship.1 == 2] <- 1
friendship.2[friendship.2 == 2] <- 1
friendship.3[friendship.3 == 2] <- 1
```

converting matrices to array

```{r}
net_size_ex2 <- nrow(friendship.1)
  
net_array <- array(c(friendship.1, friendship.2, friendship.3), 
                dim = c(net_size_ex2, net_size_ex2, 3))
```

creating network siena object

```{r}
friendship <- sienaDependent(netarray = net_array, type = "oneMode") 
```

creating behavioral siena object

```{r}
alcohol_depvar <- sienaDependent(netarray = alcohol, type = "behavior") 
```

creating gender covariate

```{r}
gender <- coCovar(sex.F) 
```

putting the whole thing together

```{r}
friend_behavior_covar <- sienaDataCreate(friendship, alcohol_depvar, gender)
```

specifying the model

```{r}
siena_effects_3wave_example <- getEffects(friend_behavior_covar) 
# starting with base model

# adding peer influence
siena_effects_3wave_example <- includeEffects(siena_effects_3wave_example, 
                                              avSim, 
                                              interaction1 = "friendship", 
                                              name = "alcohol_depvar")

# adding gender behavior effect
siena_effects_3wave_example <- includeEffects(siena_effects_3wave_example, 
                                              effFrom, 
                                              interaction1 = "gender", 
                                              name = "alcohol_depvar")

# adding effect of drinking on in-degree and out-degree
siena_effects_3wave_example <- includeEffects(siena_effects_3wave_example, 
                                              egoX, 
                                              interaction1 = "alcohol_depvar", 
                                              name = "friendship")

siena_effects_3wave_example <- includeEffects(siena_effects_3wave_example, 
                                              altX, 
                                              interaction1 = "alcohol_depvar", 
                                              name = "friendship")

# adding selection on shared behavior
siena_effects_3wave_example <- includeEffects(siena_effects_3wave_example, 
                                              simX,  
                                              interaction1 = "alcohol_depvar", 
                                              name = "friendship")

# adding transitivity
siena_effects_3wave_example <- includeEffects(siena_effects_3wave_example, 
                                              transTrip, name = "friendship")
```

estimate the model

```{r}
# set up control inputs
input_options_3wave <- sienaAlgorithmCreate(MaxDegree = c(friendship = 6),
                                            n3 = 3000, seed = 5000) 
```

```{r}
mod1_3wave_example <- siena07(input_options_3wave, 
                              data = friend_behavior_covar, 
                              effects = siena_effects_3wave_example, 
                              returnDeps = TRUE, initC = TRUE, 
                              useCluster = TRUE, nbrNodes = 2) 

mod1_3wave_example
```

checking model fit

```{r}
gof1 <- sienaGOF(mod1_3wave_example, IndegreeDistribution, 
                 verbose = TRUE, join = TRUE, varName = "friendship") 

plot(gof1)
# looks okay

gof_behaviour <- sienaGOF(mod1_3wave_example, BehaviorDistribution,
                          varName = "alcohol_depvar")
plot(gof_behaviour)
```

testing for time heterogeneity

```{r}
timetest <- sienaTimeTest(mod1_3wave_example)
summary(timetest)

# joint significance p-value is not significant; no need for a model that allows variance based on time period
```

```{r}
siena_effects_3wave_example <- includeTimeDummy(siena_effects_3wave_example, 
                                                density, 
                                                transTrip, 
                                                timeDummy = "2")
# putting in that time variance term for kicks
```

running model

```{r}
mod2_3wave_example <- siena07(input_options_3wave,
                              data = friend_behavior_covar,
                              effects = siena_effects_3wave_example, 
                              returnDeps = TRUE, 
                              initC = TRUE, useCluster = TRUE, 
                              nbrNodes = 2) 

mod2_3wave_example
```

