---
title: "coding_challenge_ergms"
format: html
---

### loading packages

```{r}
library(ergm)
library(sna)
library(car)
```

### reading in data

adjacency matrix

```{r}
url1 <- "https://github.com/JeffreyAlanSmith/Integrated_Network_Science/raw/master/data/example_schoolmat.csv"

school_mat <- read.csv(file = url1, row.names = 1) 
```

```{r}
school_mat <- as.matrix(school_mat)

school_mat[1:5, 1:5] # no friends :(
```

attributes

```{r}
url2 <- "https://github.com/JeffreyAlanSmith/Integrated_Network_Science/raw/master/data/example_schooldata.csv"

school_attributes <- read.csv(file = url2) 
```

```{r}
head(school_attributes)
```

recoding gender attribute

```{r}
school_attributes$gender_recode <- recode(school_attributes$gender,
                                          as.factor = F,
                                          "0 = 'male'; 1 = 'female'")
```

```{r}
attribute_list <- do.call(list, school_attributes)
```

### creating network

```{r}
school_net <- network(x = school_mat,
                      directed = T,
                      vertex.attr = attribute_list)

school_net
```

### basic network exploration

```{r}
plot(school_net)

# 2 isolates; not suuper dense; fair amounts of reciprocity and triadic closure (on visual observation)
```

looking at node gender 

```{r}
cols <- recode(school_attributes$gender, as.factor = F,
               "0 = 'blue'; 1 = 'pink'")
```

```{r}
plot(school_net, vertex.col = cols) 

# on biref glance, boys seem to be more central and more homophilous
```

looking at node grade

```{r}
plot(school_net, vertex.col = school_attributes$grade) 

# somewhat homophilous by individual grade; large divide between underclassmen and upperclassmen
```

looking at node degree

```{r}
outdegree <- degree(school_net, cmode = "outdegree")
indegree <- degree(school_net, cmode = "indegree")
```

```{r}
tapply(indegree, school_attributes$gender_recode, mean) 
tapply(outdegree, school_attributes$gender_recode, mean) 

# boys have higher average indegree and outdegree (gender homophily increases correlation between the 2)
```

### CUG tests

network density

```{r}
gden(school_net)
# density of .125; fairly sparse (compared to what i would expect for a school friendship network)
```

```{r}
cug.test(dat = school_net, FUN = "gden", cmode = "size") 

# also mathematically way more sparse than one would expect at random
```

reciprocity

```{r}
grecip(school_net, measure = "dyadic.nonnull")

# less than 50% reciprocity; lower than I would expect for friendship networks
```

```{r}
cug.test(dat = school_net, FUN = "grecip", 
         FUN.args = list(measure = "dyadic.nonnull"), cmode = "edges") 

# mathematically much more reciprocal than one would expect at random; but I would not expect friendship reciprocity to be random so
```

### fitting an ergm

edges only prediction

```{r}
mod_rand <- ergm(formula = school_net ~ edges) 

summary(mod_rand) 

# probability of any given tie existing is just equal to the network density 
```

```{r}
summary(school_net~edges) 
```

adding gender and grade homophily

```{r}
mod_homoph1 <- ergm(school_net ~ edges + 
                      nodematch("gender_recode") + 
                      nodematch("grade")) 

summary(mod_homoph1)

# strong homophily on grade, but not significant on gender
# a tie is 10x more likely if there is a match on gender
```

adding nodal gender and grade covariates

```{r}
mod_homoph2 <- ergm(school_net ~ edges + 
                      nodematch("gender_recode") +
                      nodematch("grade") + 
                      nodeifactor("gender_recode") + 
                      nodeofactor("gender_recode") +
                      nodeicov("grade") + 
                      nodeocov("grade"))

summary(mod_homoph2)

# men and upperclassmen tend to have more incoming and outgoing ties
```

adding reciprocity 

```{r}
set.seed(1012)

mod_homoph_mutual1 <- ergm(school_net ~ edges + 
                             nodematch("gender_recode") + 
                             nodematch("grade") + 
                             nodeifactor("gender_recode") + 
                             nodeofactor("gender_recode") + 
                             nodeicov("grade") + 
                             nodeocov("grade") +
                             mutual, 
                           control = control.ergm(MCMC.burnin = 50000,
                                                  MCMC.samplesize = 6000)) 
```

### checking model convergence

```{r}
mcmc.diagnostics(mod_homoph_mutual1, vars.per.page = 4)

# everything is pretty centered on 0, suggesting a good level of convergence
```

```{r}
summary(mod_homoph_mutual1)

# higher in/outdegreeS for men is explained by reciprocity!
```

checking reciprocity by gender

```{r}
mod_homoph_mutual1b <- ergm(school_net ~ edges + 
                              nodematch("gender_recode") +
                              nodematch("grade") + 
                              nodeicov("grade") + 
                              nodeocov("grade") +
                              mutual(by = "gender_recode"),
                            control = control.ergm(MCMC.burnin = 50000, 
                                                   MCMC.samplesize = 6000))

summary(mod_homoph_mutual1b)

# men have more mutual dyads than women
```

adding back node gender

```{r}
mod_homoph_mutual1c <- ergm(school_net ~ edges +
                              nodematch("gender_recode") + 
                              nodematch("grade") + 
                              nodeofactor("gender_recode") + 
                              nodeicov("grade") + 
                              nodeocov("grade") + 
                              mutual(by = "gender_recode"),
                            control = control.ergm(MCMC.burnin = 50000, 
                                                   MCMC.samplesize = 6000))

summary(mod_homoph_mutual1c)

# men's higher number of mutual dyads is not from higher levels of reciprocity, but from sending out more ties
```

### goodness of fit

```{r}
summary(school_net ~ esp(0:10)) 
```

```{r}
gof_mod_homoph_mutual1 <- gof(mod_homoph_mutual1, 
                              GOF = ~ distance + espartners + model,
                              control = control.gof.ergm(seed = 110))

par(mfrow = c(1, 3)) 
plot(gof_mod_homoph_mutual1)

gof_mod_homoph_mutual1

# oh no this model is not good 
```

allowing differences in homophily for each specific grade 

```{r}
mod_homoph_mutual2 <- ergm(school_net ~ edges + 
                             nodematch("gender_recode") + 
                             nodematch("grade", diff = T) + 
                             nodeifactor("gender_recode") + 
                             nodeofactor("gender_recode") + 
                             nodeicov("grade") + nodeocov("grade") + 
                             mutual)

summary(mod_homoph_mutual2)

# most homophily in grades 7 and 12 (makes sense, since they only have 1 bordering grade)
```

```{r}
gof_mod_homoph_mutual2 <- gof(mod_homoph_mutual2, 
                              GOF = ~ distance + espartners + model,
                              control = control.gof.ergm(seed = 113)) 

par(mfrow = c(1, 3)) 
plot(gof_mod_homoph_mutual2)

# model lowkey still a bad fit
```

### adding a gwesp

```{r}
mod_homoph_mutual_gwesp1 <- ergm(school_net ~ edges + 
                                   nodematch("gender_recode") + 
                                   nodematch("grade") + 
                                   nodeifactor("gender_recode") + 
                                   nodeofactor("gender_recode") + 
                                   nodeicov("grade") + nodeocov("grade") + 
                                   mutual + gwesp(decay = 1, fixed = T), 
                                 control = control.ergm(MCMLE.maxit = 5))

# this one keeps breaking :(
```
 
 let's try a different one

```{r}
mod_homoph_mutual_gwesp2 <- ergm(school_net ~ edges + 
                                   nodematch("gender_recode") +
                                   nodematch("grade") + 
                                   nodeifactor("gender_recode") + 
                                   nodeofactor("gender_recode") + 
                                   nodeicov("grade") + nodeocov("grade") +
                                   mutual + gwesp(decay = .5, fixed = T),
                                 control = control.ergm(MCMC.burnin = 50000, 
                                                    MCMC.samplesize = 7000, 
                                                    parallel = 2, 
                                                    parallel.type = "PSOCK"))

mcmc.diagnostics(mod_homoph_mutual_gwesp2, vars.per.page = 3) 
# convergence is pretty good

gof_mod_homoph_mutual_gwesp2 <- gof(mod_homoph_mutual_gwesp2, 
                                    GOF = ~ distance + espartners + model, 
                                    control = control.gof.ergm(seed = 108)) 

par(mfrow = c(1, 3))
plot(gof_mod_homoph_mutual_gwesp2) 

summary(mod_homoph_mutual_gwesp2)

# this model is looking pretty good!
```

### simulation

```{r}
net <- network.initialize(n = 32, directed = T)

net %v% "grade" <- get.vertex.attribute(school_net, "grade")

summary(school_net ~ edges + nodematch("grade")) 
```

```{r}
mod_stronghomophily <- ergm(net ~ edges + nodematch("grade"),
                            target.stats = c(124, 93)) 
mod_stronghomophily

mod_weakhomophily <- ergm(net ~ edges + nodematch("grade"), 
                          target.stats = c(124, 31))
```

```{r}
sim_strong_homophily <- simulate(mod_stronghomophily, seed = 1006) 
summary(sim_strong_homophily ~ edges + nodematch("grade")) 


sim_weak_homophily <- simulate(mod_weakhomophily, seed = 1009) 
summary(sim_weak_homophily ~ edges + nodematch("grade")) 

par(mfrow = c(1, 2))
plot(sim_weak_homophily, vertex.col = "grade", main = "Weak Homophily")
plot(sim_strong_homophily, vertex.col = "grade", main = "Strong Homophily")
```

```{r}
gtrans(sim_weak_homophily)
gtrans(sim_strong_homophily)
```

### ergms on large networks

building the network

```{r}
url3 <- "https://github.com/JeffreyAlanSmith/Integrated_Network_Science/raw/master/data/coauthorship_edgelist.txt"

coauthor_edgelist <- read.table(file = url3, header = T)
```

```{r}
url4 <- "https://github.com/JeffreyAlanSmith/Integrated_Network_Science/raw/master/data/coauthorship_attributes.txt"

coauthor_attributes <- read.table(file = url4, header = T)
```

```{r}
coauthor_attributes$prestige <- recode(coauthor_attributes$prestige,
                                       as.factor = F, 
                                       "0 = 'low prestige'; 
                                       1 = 'high prestige'")

coauthor_attributes$research_type <- recode(coauthor_attributes$research_type, 
                                            as.factor = F, 
                                            "0 = 'qualitative'; 
                                            1 = 'quantitative'; 
                                            2 = 'mixed'")
```

```{r}
coauthorship_net <- network(x = coauthor_edgelist, directed = F, 
                            vertices = coauthor_attributes) 

coauthorship_net
```

running the ergm

```{r}
mod1 <- ergm(coauthorship_net ~ edges + nodematch("research_type") +
               nodematch("prestige"))

summary(mod1)
```

```{r}
mod2 <- ergm(coauthorship_net ~ edges + nodematch("research_type") +
               nodematch("prestige") + nodefactor("research_type") + 
               nodefactor("prestige"))

summary(mod2)
```

sampling ergms that don't take a year and a half to run

```{r}
library(ergm.ego)
```

```{r}
url5 <- "https://github.com/JeffreyAlanSmith/Integrated_Network_Science/raw/master/data/coauthorship_egodat.Rdata"

load(url(description = url5))
```

```{r}
set.seed(200)
coauthorship_samp <- sample(coauthorship_egodat, 25000, replace = T)
```

```{r}
mod1_samp <- ergm.ego(coauthorship_samp ~ edges + 
                        nodematch("research_type") +  
                        nodematch("prestige"), 
                      popsize = 60098, 
                      control = control.ergm.ego(ppopsize = 25000, 
                                        ergm = control.ergm(parallel = 2)))

summary(mod1_samp)
```

### ergms on valued networks

```{r}
url6 <- "https://github.com/JeffreyAlanSmith/Integrated_Network_Science/raw/master/data/social_interaction_edgelist.txt"

social_interactions <- read.table(file = url6, header = T)


url7 <- "https://github.com/JeffreyAlanSmith/Integrated_Network_Science/raw/master/data/social_interaction_attributes.txt"

attributes <- read.table(file = url7, header = T, stringsAsFactors = F)
```

```{r}
social_net <- network(x = social_interactions, directed = F,
                      vertices = attributes)
```

looking at the network

```{r}
library(ggplot2)
library(GGally)

ggnet2(social_net, node.color = "gender", node.size = 7,
       palette = c("male" = "navy", "female" = "lightskyblue"), 
       edge.size = get.edge.attribute(social_net, "count") / 2.75, 
       edge.color = "grey80") +
  guides(size = "none")

# seemingly strong gender homophily and good mix of strong and weak ties. good amount of triadic closure?
```

ergm time

```{r}
library(ergm.count)
```

```{r}
count_mod1 <- ergm(social_net ~ sum, response = "count", 
                   reference = ~ Poisson)

summary(count_mod1)

# expect about 1 interaction (tie value of 1) for each dyad
```

adding node and matching characteristics

```{r}
count_mod2 <- ergm(social_net ~ sum + nodematch("gender", form = "sum") + 
                     nodematch("race", form = "sum") + 
                     nodefactor("gender", form = "sum") + 
                     nodefactor("race", form = "sum"), response = "count", 
                   reference = ~ Poisson, 
                   control = control.ergm(MCMC.samplesize = 5000))

summary(count_mod2)

# men interact more and there is gender hompophily; no evidence of racial homophily
```

simulation from model

```{r}
count_sim_mod2 <- simulate(count_mod2, nsim = 1)

ggnet2(count_sim_mod2, node.color = "gender", node.size = 7,
       palette = c("male" = "navy", "female" = "lightskyblue"), 
       edge.size = get.edge.attribute(count_sim_mod2, "count") / 2.75, 
       edge.color = "grey80") +
  guides(size = "none")

# looks much denser than the actual network!
```

looking at the number of valued ties

```{r}
count_sim_mod2_stats <- simulate(count_mod2, nsim = 100,
                                 monitor = ~ nonzero, output = c("stats"))

data.frame(sim = mean(count_sim_mod2_stats[, "nonzero"]), 
           empirical = summary(social_net ~ nonzero, response = "count"))

# huge overestimation
```

adding present ties as a model consideration

```{r}
count_mod3 <- ergm(social_net ~ sum + nodematch("gender", form = "sum") + 
                     nodematch("race", form = "sum") + 
                     nodefactor("gender", form = "sum") + 
                     nodefactor("race", form = "sum") + nonzero, 
                   response = "count", 
                   reference = ~ Poisson,
                   control = control.ergm(MCMC.samplesize = 5000))

summary(count_mod3)
```

transitivity in valued networks

```{r}
count_mod4 <- ergm(social_net ~ sum + nodematch("gender", form = "sum") + 
                     nodematch("race", form = "sum") + 
                     nodefactor("gender", form = "sum") + 
                     nodefactor("race", form = "sum") + nonzero + 
                     transitiveweights("min", "max", "min"), 
                   response = "count", 
                   reference = ~ Poisson, 
                   control = control.ergm(MCMC.samplesize = 5000))

summary(count_mod4)

# not a whole lot of transitivity
```

